---
Assignment 4: Linear Regression
Team Members: Vedant Naik, Rohan Joshi
output: html_document
---

### Assignment 4: Linear Regression
## Team Members: Vedant Naik, Rohan Joshi

#Problem Statement:
The price of a ticket depends, in part, on the amount of fuel consumed in the particular itinerary (a factor of distance and winds). Furthermore, prices have a tendency to increase over time. To understand airline pricing, computer a simple linear regression that models the cost of tickets for different airlines. Give a new ranking of airlines with respect to price.

The output of the Map-Reduce program is used as an input for this report.

# Airlines active in 2015

HA, EV, MQ, OO, US, B6, WN, UA, DL, NK, VX, AS, F9, AA
(Of these, NK does not have any data in 2010-2014)

# Graphs showing a linear fit of the variable to price for each airline

```{r echo=FALSE}

require(dplyr)
require(ggplot2)
library(R.utils)

# Read data folder name from command line arguments
#args <- commandArgs(trailingOnly = TRUE)
#outputFolder <- args[1]

outputFolder <- "outputPseudo"

files <- dir(outputFolder, pattern='part-r-', full.names = TRUE)
partOutputs <- lapply(files, read.csv, header=FALSE, sep="\t")
data <- bind_rows(partOutputs)
# Set headers
names(data) <- c("Carrier", "Distance", "Time", "AvgPrice")

mseDistanceLowerCount <- 0
mseTimeLowerCount <- 0

timePredMap <- new.env(hash = T, parent = emptyenv())
timePredFor <- mean(data$Time)

distancePredMap <- new.env(hash = T, parent = emptyenv())
distancePredFor <- mean(data$Distance)

for(carrier in unique(data$Carrier)){
  dfForCarrier <- data[which(data$Carrier == carrier),]
  
  lrDistance <- lm(dfForCarrier$AvgPrice ~ dfForCarrier$Distance)
  summaryDistance <- summary(lrDistance)
  mseDistance <- mean(summaryDistance$residuals^2)
  
  lrTime <- lm(dfForCarrier$AvgPrice ~ dfForCarrier$Time)
  summaryTime <- summary(lrTime)
  mseTime <- mean(summaryTime$residuals^2)
  
  if(mseTime > mseDistance){
    mseDistanceLowerCount <- mseDistanceLowerCount + 1
  } else {
    mseTimeLowerCount <- mseTimeLowerCount + 1
  }
  
  par(mfrow = c(2,1))
  
  plot(dfForCarrier$Distance, dfForCarrier$AvgPrice,ylab="Average Ticket Price",
      xlab=paste("Distance   ","Intercept :", lrDistance$coefficients[[1]], 
                 "  Slope: ", lrDistance$coefficients[[2]]))
  abline(lrDistance, col='green')
  
  plot(dfForCarrier$Time, dfForCarrier$AvgPrice,ylab="Average Ticket Price",
       xlab=paste("Distance   ","Intercept :", lrTime$coefficients[[1]], 
                  "  Slope: ", lrTime$coefficients[[2]]))
  abline(lrTime, col='red')
  
  timePredVal <- lrTime$coefficients[[2]]*timePredFor +  lrTime$coefficients[[1]]
  timePredMap[[carrier]] <- timePredVal

  distancePredVal <- lrDistance$coefficients[[2]]*distancePredFor +  lrDistance$coefficients[[1]]
  distancePredMap[[carrier]] <- distancePredVal
}

carName <- "xx"
minPrice <- 999999

finalResultList <- c()

if(mseTimeLowerCount > mseDistanceLowerCount){
  # get best price from TIME
  for(uc in unique(data$Carrier)){
    if(timePredMap[[uc]] < minPrice){
      carName <- uc
      minPrice <- timePredMap[[uc]]
    }
    
    finalResultList <- c(finalResultList, paste(timePredMap[[uc]],sep = ",",uc))
  }
  
} else {
  # get best price from DISTANCE
  for(uc in unique(data$Carrier)){
    if(distancePredMap[[uc]] < minPrice){
      carName <- uc
      minPrice <- distancePredMap[[uc]]
    }  
  }
  
  finalResultList <- c(finalResultList, paste(distancePredMap[[uc]],sep = ",",uc))
}


finalResultList <- sort(finalResultList)
finalResultList <- paste(finalResultList, sep = ",", rank(finalResultList))


cat("Cheapest carrier : ", carName)

```



# Is distance traveled or flight time a better variable?

In order to decide which model is better for this dataset, we compare the MSE (Mean Square Error) values of each model, for each carrier.
Of all the carriers seen so far `r mseTimeLowerCount` carriers have better fit for time. And `r mseDistanceLowerCount` have a better fit for
Distance.

MSE values can be calculated as follows:
```
# lrDistance <- lm(dfForCarrier$AvgPrice ~ dfForCarrier$Distance)
# summaryDistance <- summary(lrDistance)
# mseDistance <- mean(summaryDistance$residuals^2)

# lrTime <- lm(dfForCarrier$AvgPrice ~ dfForCarrier$Time)
# summaryTime <- summary(lrTime)
# mseTime <- mean(summaryTime$residuals^2)
```

We maintain a count for each variable, and increment it when the MSE for that is lower, as we iterate through all carriers
```
#  if(mseTime > mseDistance){
#    mseDistanceLowerCount <- mseDistanceLowerCount + 1
#  } else {
#    mseTimeLowerCount <- mseTimeLowerCount + 1
#  }
```

Thus we conclude the linear regression model for time has a better fit over the given data. 

# Cheapest Airline

We can use the time variable to find the cheapest airline.
Using the LR Model for time, we can get an estimated ticket prices for each airline. (Explained in detail in Analysis.)

`r cat("Cheapest carrier : ", carName) `

# ANALYSIS

We used the MSE to dertermine the best fit. In most of the cases (for most of the carriers) in this dataset, time variable gives a better fit. To automate the process of identifying which of the two variables are better, we used a naive approach of counting the lower MSEs for each carrier. (Discussed earlier.)

Once the better variable was identified, we need to find which airline is the cheapest.
For this, we used the mean value of Time, from the enitre dataset. Here, `r timePredFor`. Using this value, we predict the price for each carrier using the Linear Regression model for time.

```
# timePredVal <- lrTime$coefficients[[2]]*timePredFor +  lrTime$coefficients[[1]]
```
The slope is obainted from lrTime    $coefficients[[2]], and intercept from lrTime    $coefficients[[1]]

We store the timePredVal for each carrier in a map. Later, using this map, we find the cheapest airline, and rankings for each airline.

# Output:

The cheapest airline in this dataset: `r carName`

The rakings are given:
```{r}
for (x in strsplit(finalResultList, ",")) {
  print(paste(x[2],x[3]))
}
```


# Conclusion:

We are using MSE to determine the best fit. In this case, the time variable has a better linear regression model. Hence, we use this to predict prices
of all carriers. We are predicting based on the mean value of time from the entire dataset(here, 111.05). Of all the predicted prices(for all carriers)
F9 gives the cheapest predicted price. Hence, F9 is the cheapest.

## References:
For calculating MSE using summary of model given by lm.
http://stats.stackexchange.com/questions/107643/how-to-get-the-value-of-mean-squared-error-in-a-linear-regression-in-r
